---
title: "Group_Project_04072022"
author: "Chuck Zhou, Carl Weedman, Mark Bruno, Martin Hvezda, Patrick Holstad, and Felipe Prata"
date: "4/7/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Q&A

1. Business Problem: We are trying to predict who are "no shows". Understanding who are "no shows" will allow us to opitimize overbooking for hotels.

2. We are using a modified data set on hotel bookings where we have 119,386 rows of data with 72 columns. We are using this data to predict "no-show" customers using all available columns.

3. Number of levels for columns "agent, country, and company" presents a problem for creating dummy variables. Solved this by picking top 5-6 in each column and categorizing all others into "other".

4. "No-Show" data allows client to optimize overbooking at the hotel and maximize revenue.

5. Pursuing this project will potentially optimize revenue yield management.

## Data Exploration
```{r warning=FALSE, results=FALSE}
library(caret)
library(C50)
library(gmodels)
library(class)
library(neuralnet)
hotel <- read.csv("hotel_bookings.csv", stringsAsFactors = TRUE)
```
## Data Cleaning
### Reducing # of Levels for Agents, Company, and Country
```{r}
#sort(table(hotel$agent), decreasing = TRUE)[1:10]
hotel$newagent <- ifelse(hotel$agent == "9" | hotel$agent == "NULL" | hotel$agent == "240" | hotel$agent == "1"| hotel$agent == "14" | hotel$agent == "7" | hotel$agent == "6",  hotel$agent, "other")
#sort(table(hotel$newagent), decreasing = TRUE)
#sort(table(hotel$company), decreasing = TRUE)[1:10]
hotel$newcompany <- ifelse(hotel$company == "40" | hotel$company == "NULL" | hotel$company == "223" ,  hotel$company, "other")
#sort(table(hotel$newcompany), decreasing = TRUE)
#sort(table(hotel$country), decreasing = TRUE)[1:10]
hotel$newcountry <- ifelse(hotel$country == "PRT" | hotel$country == "GBR" | hotel$country == "FRA" | hotel$country == "ESP" | hotel$country == "DEU",  hotel$country, "other")
#sort(table(hotel$newcountry), decreasing = TRUE)[1:10]
hotel$newagent <- as.factor(hotel$newagent)
hotel$newcompany <- as.factor(hotel$newcompany)
hotel$newcountry <- as.factor(hotel$newcountry)
```

### Removing Unnecessary Columns
```{r}
hotel$reservation_status_date <- NULL
hotel$arrival_date_year <- NULL
hotel$agent <- NULL
hotel$company <- NULL
hotel$country <- NULL
hotel$reservation_status <- NULL
str(hotel)
```
## Creating Dummy Variables
```{r}
hoteldummy <- as.data.frame(model.matrix(~.-1,hotel))
hoteldummy$market_segmentUndefined <- NULL
str(hotel)
table(hotel$is_canceled) # Data is balanced
```

### Removing Unwanted Characters
```{r}
colnames(hoteldummy) <- gsub("-","",colnames(hoteldummy))
```

## Normalize and Randomize Data
```{r}
normalize <- function(x){
  (x-min(x))/(max(x)-min(x))
}
hotel_n <- as.data.frame(lapply(hoteldummy, normalize))
hotelrand <- hotel_n[sample(1:nrow(hotel_n)),]

set.seed(12345)
split <- sample(1:nrow(hotelrand), 0.5*nrow(hotelrand))
hotel_train <- hotelrand[split,]
hotel_test <- hotelrand[-split, ]
```

## Build the models

```{r cache=TRUE, warning=FALSE}
# LOGIT
#model_logit <- glm(is_canceled ~., data = hotel_train,  family = "binomial")
load("model_logit.Rdata")
logit_pred <- predict(model_logit, hotel_test, type = "response")
binary_logit <- ifelse(logit_pred >= 0.5, 1, 0)
CrossTable(binary_logit, hotel_test$is_canceled, prop.chisq = F, prop.r=F, prop.t = F, dnn=c('Actuals','Prediction'))
kappa_logit <- caret::confusionMatrix(as.factor(hotel_test$is_canceled), as.factor(binary_logit), positive = "1")$overall[2]
acc_logit <- caret::confusionMatrix(as.factor(hotel_test$is_canceled), as.factor(binary_logit), positive = "1")$overall[1]
cat('Linear Regression: ','Accuracy Ratio: ',acc_logit,'Kappa:',kappa_logit)
```


```{r cache=TRUE, warning=FALSE}
# KNN
knn_pred <- knn(train = hotel_train[, names(hotel_train) != "is_canceled"], test = hotel_test[, names(hotel_test) != "is_canceled"], cl = hotel_train$is_canceled, k = 303) 
#load("model_knn.Rdata")
CrossTable(knn_pred, hotel_test$is_canceled, prop.chisq = F, prop.r=F, prop.t = F, dnn=c('Actuals','Prediction'))
kappa_knn <- caret::confusionMatrix(as.factor(hotel_test$is_canceled), as.factor(knn_pred), positive = "1")$overall[2]
acc_knn <- caret::confusionMatrix(as.factor(hotel_test$is_canceled), as.factor(knn_pred), positive = "1")$overall[1]
cat('KNN: ','Accuracy Ratio: ',acc_knn,'Kappa:',kappa_knn)
```

```{r cache=TRUE, warning=FALSE}
# ANN
#ann <- neuralnet::neuralnet(is_canceled ~ . ,data = hotel_train)
load("model_ann.Rdata")
ann_pred <- predict(ann, hotel_test)

binary_ann <- ifelse(ann_pred >= 0.4, 1, 0)
CrossTable(hotel_test$is_canceled, binary_ann, prop.chisq = F, prop.r=F, prop.t = F, dnn=c('Actuals','Prediction'))
kappa_ann <- caret::confusionMatrix(as.factor(hotel_test$is_canceled), as.factor(binary_ann), positive = "1")$overall[2]
acc_ann <- caret::confusionMatrix(as.factor(hotel_test$is_canceled), as.factor(binary_ann), positive = "1")$overall[1]
cat('ANN: ','Accuracy Ratio: ',acc_ann,'Kappa:',kappa_ann)
```

```{r cache=TRUE, warning=FALSE}
# Decision Tree
dt <- C5.0(as.factor(is_canceled) ~. , data=hotel_train)
dt_pred <- predict(dt, hotel_test)
CrossTable(hotel_test$is_canceled, dt_pred, prop.chisq = F, prop.r=F, prop.t = F, dnn=c('Actuals','Prediction'))
kappa_dt <- caret::confusionMatrix(as.factor(hotel_test$is_canceled), as.factor(dt_pred), positive = "1")$overall[2]
acc_dt <- caret::confusionMatrix(as.factor(hotel_test$is_canceled), as.factor(dt_pred), positive = "1")$overall[1]

cat('Decision Tree: ','Accuracy Ratio: ',acc_dt,'Kappa:',kappa_dt)
```
## Build a stacked model


```{r warning= FALSE}
combined <- data.frame(binary_logit,knn_pred, binary_ann,dt_pred,hotel_test$is_canceled)
colnames(combined) <- c('binary_logit','knn_pred', 'binary_ann','dt_pred','is_canceled')
## Getting Train and Test Samples
# Selects 10000 random rows for test data
set.seed(12345)
test_set <- sample(1:nrow(combined), 0.2*nrow(combined))
hotel_test_2 <- combined[test_set,]
hotel_train_2 <- combined[-test_set, ]
# Second-Level Decision Tree
combined_model <- C5.0(as.factor(is_canceled) ~. , data=hotel_train_2)
combined_pred <- predict(combined_model, hotel_test_2)
CrossTable(hotel_test_2$is_canceled, combined_pred, prop.chisq = F, prop.r=F, prop.t = F, dnn=c('Actuals','Prediction'))
kappa_combined <- caret::confusionMatrix(as.factor(hotel_test_2$is_canceled), as.factor(combined_pred), positive = "1")$overall[2]
acc_combined <- caret::confusionMatrix(as.factor(hotel_test_2$is_canceled), as.factor(combined_pred), positive = "1")$overall[1]
plot(combined_model)
cat('Combined model: ','Accuracy Ratio: ',acc_combined,'Kappa:',kappa_combined)
```
