---
title: "Group_Project_04072022"
author: "Chuck Zhou, Carl Weedman, Mark Bruno, Martin Hvezda, Patrick Holstad, and Felipe Prata"
date: "4/7/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data Exploration
```{r}
hotel <- read.csv("hotel_bookings.csv", stringsAsFactors = TRUE)
```
## Data Cleaning
### Reducing # of Levels for Agents, Company, and Country
```{r}
#sort(table(hotel$agent), decreasing = TRUE)[1:10]
hotel$newagent <- ifelse(hotel$agent == "9" | hotel$agent == "NULL" | hotel$agent == "240" | hotel$agent == "1"| hotel$agent == "14" | hotel$agent == "7" | hotel$agent == "6",  hotel$agent, "other")
#sort(table(hotel$newagent), decreasing = TRUE)

#sort(table(hotel$company), decreasing = TRUE)[1:10]
hotel$newcompany <- ifelse(hotel$company == "40" | hotel$company == "NULL" | hotel$company == "223" ,  hotel$company, "other")
#sort(table(hotel$newcompany), decreasing = TRUE)

#sort(table(hotel$country), decreasing = TRUE)[1:10]
hotel$newcountry <- ifelse(hotel$country == "PRT" | hotel$country == "GBR" | hotel$country == "FRA" | hotel$country == "ESP" | hotel$country == "DEU",  hotel$country, "other")
#sort(table(hotel$newcountry), decreasing = TRUE)[1:10]

hotel$newagent <- as.factor(hotel$newagent)
hotel$newcompany <- as.factor(hotel$newcompany)
hotel$newcountry <- as.factor(hotel$newcountry)
```
### Removing Unnecessary Columns
```{r}
hotel$reservation_status_date <- NULL
hotel$arrival_date_year <- NULL
hotel$agent <- NULL
hotel$company <- NULL
hotel$country <- NULL
```
## Creating Dummy Variables
```{r}
hoteldummy <- as.data.frame(model.matrix(~.-1,hotel))

```
### Removing Unwanted Characters
```{r}
colnames(hoteldummy) <- gsub("-","",colnames(hoteldummy))

```

```{r}
#str(hoteldummy)
```

## Test and Train Set
```{r}
set.seed(12345)
test_set <- sample(1:nrow(hoteldummy), 0.2*nrow(hoteldummy))
hotel_test <- hoteldummy[test_set,]
hotel_train <- hoteldummy[-test_set, ]

```

## Linear Regression Model
```{r}

model1 <- lm(reservation_statusNoShow ~., data = hotel_train)
summary(model1)
```

## Project Q&A

1. Business Problem: We are trying to predict who are "no shows". Understanding who are "no shows" will allow us to opitimize overbooking for hotels.

2. We are using a modified data set on hotel bookings where we have 119,386 rows of data with 72 columns. We are using this data to predict "no-show" customers using all available columns.

3. Number of levels for columns "agent, country, and company" presents a problem for creating dummy variables. Solved this by picking top 5-6 in each column and categorizing all others into "other".

4. "No-Show" data allows client to optimize overbooking at the hotel and maximize revenue.

5. Pursuing this project will potentially optimize revenue yield management.
